# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r echo=TRUE}
library(plyr)
activity <- read.csv('activity.csv')
stepsbydate <- ddply(activity, .(date), summarize, total=sum(steps,na.rm=TRUE))
stepsbyinterval <- ddply(activity, .(interval), summarize, total=sum(steps,na.rm=TRUE), mean=mean(steps,na.rm=TRUE), median=median(steps,na.rm=TRUE))
# stepsbyinterval$interval <- sprintf("%04d",stepsbyinterval$interval)
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
hist(stepsbydate$total, main="Total Steps Taken Each Day", xlab="Total Steps Taken Each Day",col="green")
meansteps <- mean(stepsbydate$total,na.rm=TRUE)
mediansteps <- median(stepsbydate$total, na.rm=TRUE)
```
Mean steps per day is `r formatC(meansteps,format="d")`.  
Median steps per day is `r formatC(mediansteps,format="d")`.

## What is the average daily activity pattern?

```{r echo=TRUE}
# Customize x-axis; suppress plot x axis and replace with own values
plot(stepsbyinterval$interval, stepsbyinterval$mean, type="l", color="blue", xlab="5-Minute Time Intervals",xaxt="n", ylab="Average Steps per Time Interval")
axis(1,at=x<-stepsbyinterval$interval,labels=sprintf("%04d",x), cex.axis=0.7)
# axis(1,at=x<-stepsbyinterval$interval,labels=as.character(x), cex.axis=0.7)
intervalwithmaxsteps <- stepsbyinterval[which(stepsbyinterval$mean==max(stepsbyinterval$mean)),]$interval
maximumsteps <- max(stepsbyinterval$mean)
```
The 5-minute interval that contains the maximum number of average steps is `r intervalwithmaxsteps` with `r maximumsteps` steps.

## Imputing missing values
```{r echo=TRUE}

rowswithna <- activity[!complete.cases(activity),]
```
The total number of rows with NAs in the data set is `r nrow(rowswithna)`.

I shall use the mean for the 5 minute interval to replace the NA values.
```{r echo=TRUE}
# create another dataset activity2 from activity. merge with dataset containing average (mean) steps per interval, then replace NA step values with the mean.
activity2 <- activity
activity2 <- merge(activity2,stepsbyinterval,by.x="interval",sort=FALSE)
# Sort by date and interval
activity2<-activity2[with(activity2, order(date,interval)),]
activity2[!complete.cases(activity2),]$steps <- activity2[!complete.cases(activity2),]$mean
# Remove mean and median columns as they are no longer needed
activity2<-activity2[,c("steps","date","interval","total")]

```
Calculating total number of steps taken each day and calculating mean and median steps per day based on the new data set.
```{r echo=TRUE}
stepsbydate2 <- ddply(activity2, .(date), summarize, total=sum(steps,na.rm=TRUE))
stepsbyinterval2 <- ddply(activity2, .(interval), summarize, total=sum(steps,na.rm=TRUE), mean=mean(steps,na.rm=TRUE), median=median(steps,na.rm=TRUE))
hist(stepsbydate2$total, main="Total Steps Taken Each Day", xlab="Total Steps Taken Each Day",col="green")
#meansteps2 <- mean(stepsbydate2$total,na.rm=TRUE)
#mediansteps2 <- median(stepsbydate2$total, na.rm=TRUE)
meansteps2 <- mean(stepsbydate2$total)
mediansteps2 <- median(stepsbydate2$total)
```
New mean steps per day is `r formatC(meansteps2,format="d")`. Value in first step was `r meansteps` (difference is `r meansteps2-meansteps`.    
New median steps per day is `r formatC(mediansteps2,format="d")`. Value in first step was `r mediansteps` (difference is `r formatC(mediansteps2-mediansteps,format="d")`).

## Are there differences in activity patterns between weekdays and weekends?
