# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r echo=TRUE}
library(plyr)
activity <- read.csv('activity.csv')
stepsbydate <- ddply(activity, .(date), summarize, total=sum(steps,na.rm=TRUE))
stepsbyinterval <- ddply(activity, .(interval), summarize, total=sum(steps,na.rm=TRUE), mean=mean(steps,na.rm=TRUE))
stepsbyinterval$interval <- sprintf("%04d",stepsbyinterval$interval)
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
hist(stepsbydate$total, main="Total Steps Taken Each Day", xlab="Total Steps Taken Each Day",col="green")
meansteps <- mean(stepsbydate$total,na.rm=TRUE)
mediansteps <- median(stepsbydate$total, na.rm=TRUE)
```
Mean steps per day is `r meansteps`.  
Median steps per day is `r mediansteps`.

## What is the average daily activity pattern?

```{r echo=TRUE}
# Customize x-axis; suppress plot x axis and replace with own values
plot(stepsbyinterval$interval, stepsbyinterval$mean, type="l", color="blue", xlab="5-Minute Time Intervals",xaxt="n", ylab="Average Steps per Time Interval")
# axis(1,at=x<-stepsbyinterval$interval,labels=sprintf("%04d",x))
axis(1,at=x<-stepsbyinterval$interval,labels=as.character(x), cex.axis=0.7)
intervalwithmaxsteps <- stepsbyinterval[which(stepsbyinterval$mean==max(stepsbyinterval$mean)),]$interval
maximumsteps <- max(stepsbyinterval$mean)
```
The 5-minute interval that contains the maximum number of average steps is `r intervalwithmaxsteps` with `r maximumsteps` steps.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
